---
title: "Apple Inc. Stock Data Analysis"
author: "Omar Hernandez"
output:
  html_document:
    fig_width: 8
    fig_height: 4
    code_folding: hide
    highlight: tango
    toc: true
    toc_float:
      collapsed: true
# 📁 Project Folder Structure Overview

# - data/      → Contains the raw dataset (apple_stock.csv)
# - plots/     → Stores exported ggplot visuals for README & reports
# - docs/      → Final HTML report lives here for GitHub Pages
# - R/         → Reserved for utility functions (modularization)
# - launch.command → Script to knit Rmd and move HTML to docs/
# - AppleStockAnalysis.Rmd → Main analysis and forecasting notebook

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=4,
                      echo=FALSE, warning=FALSE, message=FALSE)
```

## Apple Inc. Stock Analysis
This project provides insights into Apple's stock performance by analyzing historical trading data, including trends in monthly average closing prices, the distribution of trading days across weekdays, and the applicability of the Central Limit Theorem through sampling methods.

## Apple Inc. Stock Dataset 
The Apple stock dataset captures daily trading information for Apple Inc. (AAPL), including opening and closing prices, highs, lows, adjusted closing prices, trading volume, and derived metrics like daily returns, enabling comprehensive analysis of stock performance, volatility, and market trends from 1980 to 2025.

```{r echo=TRUE}
# Load necessary libraries
library(tidyverse)
library(lubridate)
library(plotly)

# Load dataset from project root
apple_stock <- read.csv("data/apple_stock.csv", stringsAsFactors = FALSE)

# Convert Date column
apple_stock <- apple_stock %>%
  mutate(Date = as.Date(X, format = "%m/%d/%y"))

# Check for missing values
sum(is.na(apple_stock))
```

## Derive New Variables:
```{r echo = TRUE}
apple_stock <- apple_stock %>%
  mutate(DayOfWeek = weekdays(Date),
         Return = (Close - lag(Close)) / lag(Close))

print(head(apple_stock))

sum_data <- data.frame(
  Min = min(apple_stock$Return, na.rm = TRUE),
  Q1 = quantile(apple_stock$Return, 0.25, na.rm = TRUE),
  Median = median(apple_stock$Return, na.rm = TRUE),
  Mean = mean(apple_stock$Return, na.rm = TRUE),
  Q3 = quantile(apple_stock$Return, 0.75, na.rm = TRUE),
  Max = max(apple_stock$Return, na.rm = TRUE)
)
rownames(sum_data) <- c("Daily Return Summary")
sum_data
```

## Analyze Categorical and Numerical Variables
```{r echo = TRUE}
day_counts <- apple_stock %>% group_by(DayOfWeek) %>% summarise(Count = n())
p_cat <- ggplot(day_counts, aes(x = DayOfWeek, y = Count, fill = DayOfWeek)) +
  geom_bar(stat = "identity") +
  labs(title = "Number of Trading Days by Day of Week", x = "Day", y = "Count") +
  theme_minimal()
ggplotly(p_cat)

p_num <- ggplot(apple_stock, aes(x = Close)) +
  geom_histogram(bins = 30, fill = "steelblue", color = "black") +
  labs(title = "Distribution of Apple Closing Prices", x = "Closing Price", y = "Frequency") +
  theme_minimal()
ggplotly(p_num)
```

## Analyze the Relationship Between Numerical Variables
```{r echo = TRUE}
p_bi <- ggplot(apple_stock, aes(x = Open, y = Close)) +
  geom_point(alpha = 0.5, color = "darkgreen") +
  labs(title = "Scatter Plot: Open vs. Close Prices", x = "Open Price", y = "Close Price") +
  theme_minimal()
ggplotly(p_bi)
```

## Examine the Distribution of a Numerical Variable:
```{r echo = TRUE}
p_return <- ggplot(apple_stock, aes(x = Return)) +
  geom_histogram(bins = 30, fill = "coral", color = "black") +
  labs(title = "Distribution of Daily Returns", x = "Daily Return", y = "Frequency") +
  theme_minimal()
ggplotly(p_return)
```

## Central Limit Theorem Demonstration:
```{r echo = TRUE}
set.seed(123)
sample_means <- replicate(1000, {
  sample_data <- sample(apple_stock$Return, size = 30, replace = TRUE)
  mean(sample_data, na.rm = TRUE)
})
clt_data <- data.frame(SampleMean = sample_means)
p_clt <- ggplot(clt_data, aes(x = SampleMean)) +
  geom_histogram(bins = 30, fill = "lightblue", color = "black") +
  labs(title = "Sampling Distribution of Daily Return Means", x = "Sample Mean", y = "Frequency") +
  theme_minimal()
ggplotly(p_clt)

random_sample <- apple_stock %>% sample_n(500)
summary(random_sample$Return)

p_random <- ggplot(random_sample, aes(x = Return)) +
  geom_histogram(bins = 30, fill = "lightgreen", color = "black") +
  labs(title = "Histogram of Random Sample Returns", x = "Return", y = "Frequency") +
  theme_minimal()
ggplotly(p_random)

apple_stock <- apple_stock %>% mutate(Year = year(Date))
stratified_sample <- apple_stock %>% group_by(Year) %>% sample_frac(0.1)
summary(stratified_sample$Return)

p_stratified <- ggplot(stratified_sample, aes(x = Return)) +
  geom_histogram(bins = 30, fill = "lightblue", color = "black") +
  labs(title = "Histogram of Stratified Sample Returns", x = "Return", y = "Frequency") +
  theme_minimal()
ggplotly(p_stratified)
```

## Data Wrangling:
```{r echo = TRUE}
apple_stock <- apple_stock %>%
  mutate(Month = floor_date(Date, unit = "month"))

monthly_avg <- apple_stock %>%
  group_by(Month) %>%
  summarise(Avg_Close = mean(Close, na.rm = TRUE))

p_monthly <- ggplot(monthly_avg, aes(x = Month, y = Avg_Close)) +
  geom_line(color = "purple") +
  labs(title = "Monthly Average Closing Prices", x = "Month", y = "Average Close") +
  theme_minimal() +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggplotly(p_monthly)
```

